---
title: "Analysing home location"
format: html
editor: visual
---

# Clean environment

```{r}
rm(list=ls())
```

# Libraries

```{r}
library(sparklyr)    # For connecting to Spark
library(dplyr)       # For data manipulation
library(glue)
library(future)
library(future.apply)
library(sf)   # for file size detection
library(ggplot2)
library(h3)
library(tidyverse)
library(lubridate)
```

# Set working directory

## Server

```{r}
wd <- "/Volumes/DEBIAS/data/inputs"
```

# Data inputs

## Movement data

FB gridded movement data for week or month in March 2021 (excluding weekends).

```{r}
month <- FALSE # Set to FALSE if computing for week

csv_files <- list.files(path = "/Volumes/DEBIAS/data/inputs/fb/grids/movements/mar20-aug21", pattern = "\\.csv$", full.names = TRUE)

if (month == TRUE) {
# Select data for March of 2021  
csv_files <- csv_files[1029:(1029 + 31*3)]
} else {
  # Select data for March of 2021  
  csv_files <- csv_files[c(1071, 1074, 1077, 1080, 1083, 1086, 1089)]
}

data_list <- lapply(csv_files, read.csv)

# extract the date between underscores
names(data_list) <- sapply(basename(csv_files), function(x) {
  # use a regular expression to match the date between underscores
  gsub(".*_(\\d{4}-\\d{2}-\\d{2})_.*", "\\1", x)
})

# get the names (dates) from the list
dates <- as.Date(names(data_list))

# keep only weekdays (Mon–Fri)
keep_idx <- !(weekdays(dates) %in% c("Saturday", "Sunday"))

# subset the list
data_list <- data_list[keep_idx]
```

## Boundary vector data

### Local authority districts

```{r}
lad_sdf <- st_read("/Volumes/DEBIAS/data/inputs/geographies/boundaries/LAD_Dec_2021_GB_BFC_2022.gpkg") %>% 
  st_transform(crs="EPSG:4326") %>% 
  st_simplify(preserveTopology =T, # simplify boundaries
              dTolerance = 1000) %>%  # 1km
  sf::st_make_valid()  # check the geometry is valid
```

# Average gridded population

## Select the resident population (i.e. 00.00 - 8.00 UK time)

```{r}
# combine selected data frames into a single tabular dataset
combined_data <- do.call( rbind, data_list )

# filter data
selected_gridded_data <- combined_data %>% 
  dplyr::filter( country == "GB" ) #  for GB

# convert the time stamp column to POSIXct date-time format
selected_gridded_data$timestamp <- as.POSIXct(selected_gridded_data$date_time, format = "%Y-%m-%d %H:%M")

# filter night-time population: 0 AM pacific time is 8 AM UK time
selected_gridded_data <- subset(selected_gridded_data, format(selected_gridded_data$timestamp, "%H:%M") == "00:00")
```

## Approach 1: averaging temporally and then aggregating spatially

### Average temporally

```{r}
# group by spatial unit and calculate average over time
average_gridded_mov_df <- selected_gridded_data  %>%
  mutate(coords = str_extract(geometry, "(?<=\\().*(?=\\))")) %>%
  separate(coords, into = c("start_coord", "end_coord"), sep = ", ")  %>%
  group_by( start_coord, end_coord ) %>%
  summarise(avg_fb_mov = mean(n_crisis, na.rm = TRUE), .groups = "drop")
```

### Aggregate spatially

```{r}
average_gridded_mov_sdf <- average_gridded_mov_df %>%
  separate(start_coord, into = c("start_long", "start_lat"), sep = " ", convert = TRUE) %>%
  st_as_sf(coords = c("start_long", "start_lat"),
           crs = 4326)

average_lad_start_mov <- st_join(lad_sdf,
                    average_gridded_mov_sdf,
                    join = st_contains) %>%
  select("LAD21CD", "end_coord", "avg_fb_mov") %>%
  rename("LAD21CD_start" = "LAD21CD") %>%
  st_drop_geometry()

average_gridded_mov_sdf <- average_lad_start_mov %>%
  separate(end_coord, into = c("end_long", "end_lat"), sep = " ", convert = TRUE) %>%
  filter(!is.na(end_long)) %>%
  filter(!is.na(end_lat)) %>%
  st_as_sf(coords = c("end_long", "end_lat"),
           crs = 4326)

mov_lad_tts <- st_join(lad_sdf,
                    average_gridded_mov_sdf,
                    join = st_contains) %>%
  select("LAD21CD", "LAD21CD_start", "avg_fb_mov") %>%
  rename("LAD21CD_end" = "LAD21CD") %>%
  st_drop_geometry() %>%
  group_by( LAD21CD_start, LAD21CD_end ) %>%
  summarise(avg_fb_mov = sum(avg_fb_mov, na.rm = FALSE), .groups = "drop")

if (month == TRUE) {
  write.csv(mov_lad_tts, "/Volumes/DEBIAS/data/inputs/fb/flows/tts/htw-lad-tts-month.csv")
} else {
  write.csv(mov_lad_tts, "/Volumes/DEBIAS/data/inputs/fb/flows/tts/htw-lad-tts.csv")
}

```

Checking the share of grids with average movement counts. *How many rows are NAs?*

```{r}
# original input data
 (sum(is.na(selected_gridded_data$n_crisis)) / nrow(selected_gridded_data)) * 100
# aggregate data
 (sum(is.na(mov_lad_tts$avg_fb_mov)) / nrow(mov_lad_tts)) * 100
```

## Apporach 2: aggregating spatially and then averaging temporally

### Aggregate spatially

```{r}
aggregate_gridded_start_sdf <- selected_gridded_data %>% 
  mutate(coords = str_extract(geometry, "(?<=\\().*(?=\\))")) %>% 
  separate(coords, into = c("start_coord", "end_coord"), sep = ", ") %>% 
  separate(start_coord, into = c("start_long", "start_lat"), sep = " ", convert = TRUE) %>%
  st_as_sf(coords = c("start_long", "start_lat"), crs = 4326)

aggregate_lad_start_df <- st_join(select(lad_sdf, "LAD21CD", "SHAPE"), 
                                  aggregate_gridded_start_sdf, 
                                  join = st_contains) %>% 
  rename("LAD21CD_start" = "LAD21CD") %>%
  st_drop_geometry()

aggregate_lad_sdf <- aggregate_lad_start_df %>%
  separate(end_coord, into = c("end_long", "end_lat"), sep = " ", convert = TRUE) %>%
  filter(!is.na(end_long)) %>%
  filter(!is.na(end_lat)) %>%
  st_as_sf(coords = c("end_long", "end_lat"),
           crs = 4326)

aggregate_lad_df <- st_join(select(lad_sdf, "LAD21CD", "SHAPE"), 
                                  aggregate_lad_sdf, 
                                  join = st_contains) %>% 
  rename("LAD21CD_end" = "LAD21CD") %>%
  st_drop_geometry() %>%
  select(LAD21CD_start, LAD21CD_end, date_time, n_crisis) %>%
  group_by( LAD21CD_start, LAD21CD_end, date_time ) %>%
  summarise(n_crisis_LAD = sum(n_crisis, na.rm = FALSE), .groups = "drop")

```

### Average temporally

```{r}
mov_lad_stt <- aggregate_lad_df %>%
  group_by( LAD21CD_start, LAD21CD_end ) %>%
  summarise(avg_fb_mov = mean(n_crisis_LAD, na.rm = TRUE), .groups = "drop")

if (month == TRUE) {
  write.csv(mov_lad_stt, "/Volumes/DEBIAS/data/inputs/fb/flows/stt/htw-lad-stt-month.csv")
} else {
  write.csv(mov_lad_stt, "/Volumes/DEBIAS/data/inputs/fb/flows/stt/htw-lad-stt.csv")
}
```

Checking the share of grids with average movement counts. *How many rows are NAs?*

```{r}
# original input data
 (sum(is.na(selected_gridded_data$n_crisis)) / nrow(selected_gridded_data)) * 100
# aggregate data
 (sum(is.na(mov_lad_stt$avg_fb_mov)) / nrow(mov_lad_stt)) * 100
```

## Test consistency in approaches

```{r}
joined_df <- mov_lad_tts %>%
  inner_join(mov_lad_stt, by = c("LAD21CD_start", "LAD21CD_end")) %>%
  rename("avg_fb_mov_tts" = "avg_fb_mov.x",
         "avg_fb_mov_stt" = "avg_fb_mov.y") 

plot(joined_df$avg_fb_mov_tts, joined_df$avg_fb_mov_stt)
```
